---
title: "Practice Activity 4: Tidy Data with tidyr"
format: html
execute: 
  echo: true
  eval: false
---

## Setup

Today you will be tidying untidy data to explore the relationship between
countries of the world and military spending.

```{r packages}
library(readxl) 
library(tidyverse)
```

## Data for Today's Activity

The SIPRI (Stockholm International Peace Research Institute) Military
Expenditure Database is an open source dataset with contains consistent time
series on the military spending of countries for the period 1949--2018. The
database is updated annually, which may include updates to data for any of the
years included in the database.

Military expenditure in local currency at current prices is presented according
to both the financial year of each country and according to calendar year,
calculated on the assumption that, where financial years do not correspond to
calendar years, spending is distributed evenly through the year. Figures in
constant (2017) and current US \$, as a share of GDP and per capita are
presented according to calendar year. The availability of data varies
considerably by country, but for a majority of countries that were independent
at the time, data is available from at least the late 1950s. Estimates for
regional military expenditure have been extended backwards depending on
availability of data for countries in the region, but no estimates for total
world military expenditure are available before 1988 due to the lack of data for
the Soviet Union.

SIPRI military expenditure data is based on open sources only.

## Data Inspection

**Download the data [here](https://app.box.com/s/txssesmg6xi28ttf3kfrrfmivfegdkmb). Open the Excel file and inspect the file.**

First, you should notice that there are ten different sheets included in the
dataset. We are interested in the sheet labeled "Share of Govt. spending" which
contains information about the share of government spending that is allocated to
military spending.

Next, you'll notice that there are notes about the dataset in the first six
rows. Ugh! Also notice that the last six rows are footnotes about the dataset.
**Ugh**!

Rather than copying this one sheet into a new Excel file and deleting the first
and last six rows, let's learn something new about the `read_xlsx()` function!

## Data Import

The `read_xlsx()` function has a `sheet` argument, where you specify the name of
the sheet that you want to use.

**Note:** The name must be passed in as a string (in quotations)!

The `read_xlsx()` function also has a `skip` argument, where you specify the
number of rows you want to be skipped *before* reading in the data.

Finally, `read_xlsx()` also has a `n_max` argument, where you specify the
maximum number of rows of data to read in.

1. Modify the code below to read the military expenditures data into your
work space. You will need to modify the path to access the data. 

```{r data}
military <- read_xlsx(here::here("tidyr", 
                                 "gov_spending_per_capita.xlsx"), 
                      sheet = ____, 
                      skip = ____, 
                      nmax = ____)
```

## Data Cleaning

In addition to `NA`s, missing or unavailable values were coded two ways.

2. Find these two methods and write the code to replace these values with
NAs. Save the mutated dataset into a new object named `military_clean`.

Because of the use of characters to mark missing values, all of the columns 1988
through 2019 were read in a characters.

3. Mutate these columns to all be numeric data types, instead of a character 
data type. Save these changes into an updated version of `military_clean`.

If you give the `Country` column a look, you'll see that there are names of
continents **and** regions included. These names are only included to make it
simpler to find countries, as they contain no data.

Luckily for us, these region names were also stored in the "Regional totals"
sheet. We can us the `Region` column of this dataset to filter out the names we
don't want.

Run the code below to read in the "Regional totals" dataset, making any
necessary modifications to the path.

```{r regions-data}
cont_region <- read_xlsx(here::here("tidyr",
                                    "gov_spending_per_capita.xlsx"), 
                      sheet = "Regional totals", 
                      skip = 14) |> 
  filter(Region != "World total (including Iraq)", 
         Region != "World total (excluding Iraq)")
```

If you think about `filter()`ing data, so that only certain values of a variable
are retained, we should remember our friendly `%in%` function! However, if you 
think about the code below, you should notice that this code retains only the
values we wanted to remove!

```
military_clean |> 
  filter(Country %in% cont_region$Region)
```

Unfortunately, R doesn't come with a built-in `!%in%` function. However, a
clever way to filter out observations you don't want is with a join. A tool
tailored just for this scenario is the `anti_join()` function. This function
will return all of the rows of one dataset **without** a match in another
dataset.

4. Use the `anti_join()` function to filter out the `Country` values we don't
want in the `military_clean` dataset. The `by` argument needs to be filled with
the name(s) of the variables that the two datasets should be joined with.

**Hint:** To join by different variables on `x` and `y`, use a named vector. For
example, `by = c("a" = "b")` will match `x$a` to `y$b`.

### Part One Answer

What regions were not removed from the `military_clean` dataset? Use their
correct capitalization **and** punctuation!

## Data Organization

The comparison I am interested in looking at the military expenditures across
every year in the data. Something like this:

<https://app.box.com/embed/s/ed74ynmti87pk8oats4jdbhwb5rqf7cm?sortColumn=date&view=list>

Unfortunately, this requires that every year is included in one column!

To tidy a dataset like this, we need to pivot the columns of years from wide
format to long format. To do this process we need three parameters:

-   The set of columns that represent values, not variables. In these data,
those are all the columns from `1988` to `2018`.

-   The name of the variable that should be created to move these columns into.
In these data, this could be `"Year"`.

-   The name of the variable that should be created to move these column's
values into. In these data, this could be labeled `"Spending"`.

Each of these pieces form the three required arguments to the `pivot_longer()`
function.

5. Pivot the cleaned up `military` dataset to a "longer" orientation. Save this
new "long" version as a new dataset (**do not** overwrite your cleaned up dataset)!

## Data Visualization Exploration

Now that we've transformed the data, let's create a plot to explore the military
spending across the years.

6. Create side-by-side boxplots of the military spending for each year.

**Hint:** Place the `Year` variable on an axis that makes it easier to read the
labels!

### Part Two Answer

What year was the second largest military expenditure? What country had this
expenditure?

**Bonus**: What is the reason for this large expenditure?
