---
title: "Practice Activity: Regression"
output: rmdformats::readthedown

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(include = FALSE, eval = FALSE)
```

```{r}
altj <- c(0, 1, 1, 1, 0, 0, 1, 1)
```

```{r}
dat <- read_csv(here::here("8_Regression_Testing", "PA-regression", "data_2.csv"))

new_dat <- dat %>%
  select(-z) %>%
  mutate(
    y = y - mean(y),
    weight_before = x,
    weight_after = 1.2*x + y/5
  )

new_dat %>%
  ggplot(aes(x = weight_before, y = weight_after)) +
  geom_point()

mod <- lm(weight_after ~ weight_before, data = new_dat)

new_dat %>%
  modelr::add_predictions(mod) %>%
  ggplot(aes(x = weight_before, y = weight_after - pred)) +
  geom_point()

new_dat %>%
  select(weight_before, weight_after) %>%
  write_csv(here::here("8_Regression_Testing", "PA-regression", "mystery_animal.csv"))
```

The dataset "mystery_animal.csv" contains the weights of a particular animal before and after a year of eating only roasted duck.
 

Compute a linear model to determine if the Duck Diet caused the animal to gain weight, lose weight, or neither.
 

Plot the residuals (observed value minus predicted value) of this model versus
the x-variable.  The function `add_predictions()` from the "modelr" package may be useful to you.
