---
title: "Practice Activity 2: Objects and subsetting"
author: "Stat 331"
output: rmdformats::readthedown
---

Today you will be creating and manipulating vectors, lists, and data frames to uncover a top secret message.

## Dealing with errors

As you work through this PA, you will encounter some code that does not work as you want it to.  Don't despair!  Errors (when R is unable to run your code) and bugs (when the code doesn't do what you hoped) are a natural part of coding.  Even the best of the best deal with these problems regularly - learning to track down the issue is a skill that you can learn and practice.

### Advice for dealing with errors

**Errors can be sneaky - check results often** 

If a chunk of code runs smoothly without giving you any error or warnings this does not necessarily mean it accomplished the desired task.

It is a good habit to check the results of your code every time you finish a task.

* In the text before the code chunk, make sure to briefly state what the point of the chunk is.  This will remind later readers - which might be your future self! - what the desired output is.

* If you created a new object, take a look at it, either by clicking its name in your *Environment* tab or by typing its name into the console.  Make sure it looks about how you expect.

* If you created or updated a data frame, make sure your edits did what you hoped.  Use the *Environment* or the `head()` function to investigate your changes.


**Two heads are better than one**

It can be hard to spot bugs in code that you wrote.  

Work with those around you - if something goes wrong, ask a friend to take a peek at your code and see if any glaring errors (like *syntax error*) pop out.

**Explain your code out loud**

The best way to troubleshoot a sneaky but is to explain out loud each step of your code, and what you hoped to accomplish.

If you are alone, try [Rubber Duck Debugging](https://en.wikipedia.org/wiki/Rubber_duck_debugging)!

**Google is your friend**

The whole of the internet is at your disposal!  Use it early, use it often.

Some tricks:

* Copy-paste the exact error message into Google. Chances are, somebody else had a similar problem and got a similar message.

* Include package names in your search terms.  For example, "bar plot in ggplot" is a better search than "bar plot in R".


### Part One: Data import and cleaning

This section will clean today's dataset, so that you can use it more easily in Part Two.  

First, we declare our package dependencies and load the data.

(Note that the data loading function `read_csv` will give you an outpouring of helpful information about the dataset.  If you do not see the word "error", there is nothing to be concerned about.)

```{r setup, message = FALSE, warning = FALSE}
library(tidyverse)

colleges <- read_csv("https://www.dropbox.com/s/bt5hvctdevhbq6j/colleges.csv?dl=1")
```


Now we will clean the data.  Alas, each of the R chunks in this section will cause an error and/or do the desired task incorrectly.  (Even the chunks that run without error are not correct!) You will need to find the mistake, and correct it, to complete the intended action. 


There are too many variables in this dataset.  We don't need all of them. Narrow your dataset down to only:

Name of the institution
City, State, and ZIP code of the institution
The Admissions Rate
The average SAT score
The number of undergraduate students
The in and out of state tuitions
Whether the school is public or private
The "REGION" variable.

```{r, eval = FALSE}
colleges_clean <- colleges %>
  select(INSTNM, CITY, STABBR, ZIP, CONTROL, ADM_RATE, SAT_AVG, TUITIONFEE_IN, TUITIONFEE_OUT, UGDS, REGION) 
```

Drop the schools that are private and for-profit (category 3).

```{r, eval = FALSE}
colleges_clean <- colleges_clean %>%
  filter(CONTROL == 1, CONTROL == 2) 
```

Adjust the appropriate variables to be numeric.

```{r, eval = FALSE}
colleges_clean <- colleges_clean %>%
  mutate(
    TUITIONFEE_IN = numeric(TUITIONFEE_IN),
    TUITIONFEE_OUT = numeric(TUITIONFEE_OUT),
    SAT_AVG = numeric(SAT_AVG),
    ADM_RATE = numeric(ADM_RATE)
    ) 
```

Adjust the appropriate variables to be factors.

```{r, eval = FALSE}
colleges_clean %>% mutate(
    CONTROL = as.factor(CONTROL),
    REGION = as.factor(REGION)
)
```

Create a new variable called TUITION_DIFF which contains the difference between in and out of state costs.

```{r, eval = FALSE}
colleges_clean <- colleges_clean %>% 
    TUITION_DIFF = TUITIONFEE_OUT - TUITIONFEE_IN
```


Drop all the rows with missing data.  
(This is not always a great idea!  Usually, even if *some* of the information is missing, we don't want to throw out the entire row.  This time, however, we'll be lazy.)

```{r, eval = FALSE}
colleges_clean <- colleges_clean %>% drop.na()
```



Lastly, notice that each of these steps started with 

```{r, eval = FALSE}
colleges_clean <- colleges_clean %>% ...
```

That is pretty redundant!  Instead, we could perform all these tasks as one long "pipeline".  Combine your (fixed) code chunks into a single chunk that cleans the data.

```{r, include = FALSE}
colleges_clean <- colleges %>%
  select(INSTNM, CITY, STABBR, ZIP, CONTROL, ADM_RATE, SAT_AVG, TUITIONFEE_IN, TUITIONFEE_OUT, UGDS, REGION) %>%
  filter(CONTROL != 3) %>%
  mutate(
    TUITIONFEE_IN = as.numeric(TUITIONFEE_IN),
    TUITIONFEE_OUT = as.numeric(TUITIONFEE_OUT),
    SAT_AVG = as.numeric(SAT_AVG),
    ADM_RATE = as.numeric(ADM_RATE),
    CONTROL = as.factor(CONTROL),
    TUITION_DIFF = TUITIONFEE_OUT - TUITIONFEE_IN,
    REGION = as.factor(REGION)
    ) %>%
    drop_na()
```



## Part Two:  Identify the mystery college

Wow!  Your best friend Ephelia has been accepted to the college of her dreams!  Unfortunately, Ephelia is a very mysterious person, and she won't tell you directly which college this is.  You'll have to use her clues to figure out which school is her dream school.


Clues:

1. This college is located in Region 1.

2. This college's admission rate is in the first quartile for the region.

3. This college charges the same for in- and out-of-state tuition.

4. The average SAT score of this college is an odd number.

5. This college is NOT in New Hampshire or in the city of Boston.

6. More than 3,000 people apply to this college every year.  (Assume the size of the freshman class is 1/4 of the undergraduate population.)

7. Dr. Bodwin did not attend this college.

8. Of the two options remaining at this step, Ephelia will attend the cheaper one.

```{r, include = FALSE}
colleges_clean %>%
  filter(REGION == 1,
         ADM_RATE < 0.4789,
         TUITION_DIFF == 0) %>%
  mutate(
    SAT_even = SAT_AVG %% 2 != 0
  ) %>%
  filter(SAT_even,
         CITY != "Boston",
         STABBR != "NH") %>%
  mutate(
    app = as.numeric(UGDS)/4/ADM_RATE
  ) %>%
  filter(app > 3000)
```

