---
title: "decodeR:  Matrix Algebra"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)  #ignore this
```

## Introduction

You will be asked to perform a series of calculations on a matrix of numbers.  If you follow all the steps successfully, your end result will be a matrix of pixel intensities for an image - that is, if you plot your matrix (I have supplied the code for that), you will see a black and white photograph.

## Setup 

Run the following code chunk to load up your matrix (M), and to generate some random matrices and vectors that we will use along the way.

```{r}
### DO NOT CHANGE ANYTHING IN THIS CODE CHUNK!!!

M = read.csv("./Matrix_Algebra/img.txt", header = FALSE)
M = as.matrix(M)


n = nrow(M)
p = ncol(M)

set.seed(123)
A <- diag(rnorm(p))
set.seed(456)
B <- tcrossprod(as.matrix(array(rnorm(25), c(5,5))))
set.seed(789)
x <- abs(rnorm(p))
set.seed(101)
y <- rnorm(n)
set.seed(112)
H <- as.matrix(array(rnorm(n*p), c(n,p)))

```

## Hints

### Hint 1

When you are asked to do something to the matrix M, make sure you save the updated matrix, not just perform the calculation and print it out. **At each step, the matrix M should retain its original dimensions - this is a good way to check your progress.**  For example, let's do step 1 together:

1. Matrix-multiply M by A.

```{r, eval = FALSE}
# Correct step:
M <- M %*% A

# This will perform the calculation, but not update M
# M %*% A

# This will result in an error because the dimensions do not match
#  A %*% M
```

### Hint 2

When you are doing matrix multiplications, it is typically harder to make errors, because R will not perform the calculation if the dimensions don't line up right.

When you are dealing with a matrix or vector and a scalar, it is also harder to make errors, because there are not dimensions to worry about with the scalar.

However, when you are dealing with a matrix and vector or two vectors, sometimes you can do the wrong thing and get no errors.

For example, let's look at the **wrong way** to do Step 2:


2. Multiply every row of M elementwise by x. 

```{r}
M_wrong <- M*x
dim(M)
dim(M_wrong)
```

What happened?  Well, R's default when given ordinary multiplication (`*`) on a matrix and a vector is to start multiplying down the **columns** of the matrix, not the rows.  This is because we consider all vectors, by default, to be **column vectors**.  

Here is a simpler example:

```{r}
Mat <- rbind(
  c(1,2),
  c(3,4)
)

vec <- c(1, 100)

Mat * vec
```

One solution to this problem (there are many ways to approach it!) is to be creative with transposing your matrix:

```{r}
new_mat <- t(Mat)*vec
new_mat
new_mat <- t(new_mat)
new_mat
```

### Hint 3

This process of unveiling a photo will only work if you do each step, in order, exactly once.  If you start to get thrown off, simply re-load the matrix `M` by re-running the code, and run each step again.

## Your turn

Before you finalize your process, delete everything above "Your Turn", except for the setup code chunk, so that you are not duplicating steps 1 and 2.

Perform the following steps (we have already done 1 and 2) to discover the secret picture!


1. Matrix-multiply M by A.


2. Multiply every row of M elementwise by x. 


3. Make a diagonal matrix called Y whose diagonal values are the elements of the vector y.  Matrix-multiply the inverse of Y by M.


4.  Calculate the determinant of B.  Add this scalar to every element of M.


5. Use `eigen()` to find the eigenvalues and eigenvectors of B.  Find the sum of the third eigenvector of B. Multiply every element of M by this scalar.


6. Multiply each element of M by the corresponding element of the matrix formed by (xy')'.


7. Find the singular value decomposition H = UDV' using `svd()`.  Multiply each element of M by each element of 100*U. 


8. Calculate the sum of the singular values of H.  Divide *every* element of M by this scalar.


## The results

Run the following code to plot your result.  If it does not look like a real photographic image, something has gone wrong - go back and check your steps.

```{r}
image(M, col = gray((1:100)/100), asp = 1, axes = FALSE)
```
















